Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-10-27 short.tex
very fancy table

The constrained optimization problem in
Eq.~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%Describing Sketch's program synthesis
%algorithm is beyond the scope of this paper; see~\cite{solar2008program}.
At a high level, Sketch takes as input a space of programs, along with
a specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a minimum-cost
program satisfying the specification.  Sketch requires a
 \emph{finite program space}, which here means that the depth of the
program syntax tree is bounded (we set the bound to 3),
but comes with the guarantee that it will
always eventually find the globally optimal (depth-bounded) solution.
In exchange for this optimality guarantee
it comes with no guarantees
on runtime.
For our domain synthesis times vary from minutes to hours,
with 27\% of the drawings timing out the synthesizer after 1 hour.
See Fig.~\ref{exampleSynthesisResults} for examples of the kinds of programs our system recovers.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Section~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving 58\% of problems within a minute.

@@

\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(2,8)
Rectangle(6,9, 7,10)
Circle(8,8)
Rectangle(6,12, 7,13)
Rectangle(3,9, 4,10)
\end{lstlisting}\small\emph{... etc. ...; 9 lines}
  \end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
reflect(y=8)
 for(i<3)
  if(i>0)
   rectangle(3*i-1,2,3*i,3)
  circle(3*i+1,3*i+1)
\end{lstlisting}
\end{minipage}&$\frac{9}{5} = 1.8\text{x}$ %\\\bottomrule
  \end{tabular}
\caption{Example drawings (left), their ground truth trace sets (middle left), and programs synthesized from these traces (middle right). Compared to the trace the programs are more compressive (right: programs have fewer lines than traces) and automatically group together related drawing commands. Note the nested loops in the Ising model, special case conditionals for the HMM, combination of symmetry and iteration in the bottom figure, and affine transformations in the top figure, and the complex program in the second figure to bottom.}\label{exampleSynthesisResults}
%  \vspace{-0.7cm}
  \end{figure}

@@

\begin{table}[h]\centering
  \caption{Parameterization of different ways of posing the program synthesis problem. The policy learns to choose parameters likely to quickly yield a minimal cost program. We slightly abuse notation by writing $\sigma $ to mean an assignment to each of these parameters (so $\sigma $ assumes one of 24 different values) and to mean the set of programs selected by that parameterization (so $\sigma \subseteq \text{DSL}$)}\label{policyOutput}
  \begin{tabular}{lll}\toprule
  Parameter&Description&Range\\\midrule
  Loops?&Is the program allowed to loop?&$\{\text{True},\text{False}\}$\\
  Reflects?&Is the program allowed to have reflections?&$\{\text{True},\text{False}\}$\\
  Incremental?&Solve the problem piece-by-piece or all at once?&$\{\text{True},\text{False}\}$\\
  Maximum depth& Bound on the depth of the program syntax tree&$\{1,2,3\}$
  \\\bottomrule
  \end{tabular}
  \end{table}

Given a search policy, what is the best way of using it to quickly find minimum cost programs?
We use a \emph{bias-optimal search algorithm} (see~\cite{schmidhuber2004optimal}):

@@

We measure the similarity between two drawings by extracting features
of the lowest-cost programs that describe them. Our features are counts of the number of times that different components in the
DSL were used (Tbl.~\ref{DSL}).
We  then find drawings which are either close together or far apart in program feature space.
One could use many
alternative similarity metrics between drawings which would capture pixel-level similarities while missing high-level geometric similarities.
We used our learned distance metric between trace sets, $L_{\text{learned}}(\cdot|\cdot)$,
to find drawings that are either close together or far apart according to the learned
distance metric over images.
Tbl.~\ref{similarityMatrix} illustrates the kinds of drawings that these different metrics put closely together,
while the supplement shows low dimensional projections of
program and image features.

\newcommand{\similaritySize}{1cm}
\newcommand{\similarityArrowSize}{0.35cm}
\begin{table}[t]
\centering  \caption{Pairs of images either close together or far apart in different features spaces.  The symbol $\leftrightarrow$ points to the compared images. Program features base captures abstract notions like symmetry and repetition. Distance metric over images is $L_{\text{learned}}(\cdot|\cdot)$ (see Section~\ref{generalizingTheHandDrawings}).}\label{similarityMatrix}
  \begin{tabular}{lll}
    &\textbf{Close in program space}&\textbf{Far apart in program space}\\\\
    \textbf{Close in image space}&\begin{tabular}{l}
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-20-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-21-trim.png}\end{minipage}\\\\
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-50-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-2-trim.png}\end{minipage}
      \end{tabular} &
    \begin{tabular}{l}
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-22-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-77-trim.png}\end{minipage}\\\\
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-64-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-97-trim.png}\end{minipage}
      \end{tabular}&\\
    \textbf{Far apart in image space}&
    \begin{tabular}{l}
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-52-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-79-trim.png}\end{minipage}\\\\
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-16-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-17-trim.png}\end{minipage}
    \end{tabular}&
    \begin{tabular}{l}
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-75-extra-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-89-trim.png}\end{minipage}\\\\
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-3-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-39-trim.png}\end{minipage}
      \end{tabular}
  \end{tabular}
  \end{table}

@@

\subsection{Extrapolating figures}
Having access to the source code of a graphics program facilitates coherent, high-level edits to the figure generated by that program. 
For example,
we could change all of the circles to squares or make all of the lines be dashed.
We \textbf{extrapolate} figures
by increasing the number of times that loops are executed.
Extrapolating repetitive visuals patterns comes naturally to humans,
and building this ability into an application is practical:
imagine hand drawing a repetitive graphical model structure
and having our system automatically induce and extend the pattern.
Fig.~\ref{extrapolationFigure} shows extrapolations produced by our system.
\begin{figure}[H]
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix1.png}
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix2.png}
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix3.png}  
  \caption{Top, white: hand drawings. Bottom, black: extrapolations produced by
running loops for extra iterations.}\label{extrapolationFigure}
  \end{figure}
%\input{extrapolations.tex}

>>>>

The constrained optimization problem in
Eq.~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%At a high level,
Sketch takes as input a space of programs, along with
a specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a minimum-cost
program satisfying the specification.  Sketch requires a
 \emph{finite program space}, which here means that the depth of the
program syntax tree is bounded (we set the bound to 3),
but has the guarantee that it 
always eventually finds the globally optimal solution.
In exchange for this optimality guarantee
it comes with no guarantees
on runtime.
For our domain synthesis times vary from minutes to hours,
with 27\% of the drawings timing out the synthesizer after 1 hour.
Fig.~\ref{exampleSynthesisResults} shows programs recovered by our system.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Sec.~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving 58\% of problems within a minute.

@@

\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(2,8)
Rectangle(6,9, 7,10)
Circle(8,8)
Rectangle(6,12, 7,13)
Rectangle(3,9, 4,10)
\end{lstlisting}\small\emph{... etc. ...; 9 lines}
  \end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
reflect(y=8)
 for(i<3)
  if(i>0)
   rectangle(3*i-1,2,3*i,3)
  circle(3*i+1,3*i+1)
\end{lstlisting}
\end{minipage}&$\frac{9}{5} = 1.8\text{x}$ \\\bottomrule
  \end{tabular}
\caption{Example drawings (left), their ground truth trace sets (middle left), and programs synthesized from these traces (middle right). Compared to the trace the programs are more compressive (right: programs have fewer lines than traces) and automatically group together related drawing commands. Note the nested loops in the Ising model, special case conditionals for the HMM, combination of symmetry and iteration in the bottom figure,  affine transformations in the top figure, and the complicated program in the second figure to bottom.}\label{exampleSynthesisResults}
  \end{figure}

@@

\begin{table}[h]\centering
  \caption{Parameterization of different ways of posing the program synthesis problem. The policy learns to choose parameters likely to quickly yield a minimal cost program. We slightly abuse notation by writing $\sigma $ to mean an assignment to each of these parameters (so $\sigma $ assumes one of 24 different values) and to mean the set of programs selected by that parameterization (so $\sigma \subseteq \text{DSL}$)}\label{policyOutput}
  \begin{tabular}{lll}\toprule
  Parameter&Description&Range\\\midrule
  Loops?&Is the program allowed to loop?&$\{\text{True},\text{False}\}$\\
  Reflects?&Is the program allowed to have reflections?&$\{\text{True},\text{False}\}$\\
  Incremental?&Solve the problem piece-by-piece or all at once?&$\{\text{True},\text{False}\}$\\
  Maximum depth& Bound on the depth of the program syntax tree&$\{1,2,3\}$
  \\\bottomrule
  \end{tabular}
  \end{table}
Given a search policy, what is the best way of using it to quickly find minimum cost programs?
We use a \emph{bias-optimal search algorithm}~\citep{schmidhuber2004optimal}:

@@

We measure the similarity between two drawings by extracting features
of the lowest-cost programs that describe them. Our features are counts of the number of times that different components in the
DSL were used (Tbl.~\ref{DSL}).
We  then find drawings which are either close together or far apart in program feature space.
One could use many
alternative similarity metrics between drawings which would capture pixel-level similarities while missing high-level geometric similarities.
We used our learned distance metric between trace sets, $L_{\text{learned}}(\cdot|\cdot)$,
to find drawings that are either close together or far apart according to the learned
distance metric over images.
Fig.~\ref{similarityMatrix} illustrates the kinds of drawings that these different metrics put closely together,
while the supplement shows low dimensional projections of
program and image features.

\newcommand{\similaritySize}{1cm}
\newcommand{\similarityArrowSize}{0.35cm}
\begin{figure}[t]
  \begin{tabular}{c|c|c|}
    \multicolumn{1}{c}{}    &\multicolumn{1}{c}{\textbf{Close in program space}}&\multicolumn{1}{c}{\textbf{Far apart in program space}}\\\cline{2-3}
    \multicolumn{1}{c|}{}&&\\
    \multicolumn{1}{c|}{    \begin{tabular}{l}\textbf{Close in }\\ \textbf{image space} \end{tabular}}    &
    \begin{tabular}{l}
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-20-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-21-trim.png}\end{minipage}\\\\
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-50-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-2-trim.png}\end{minipage}
    \end{tabular} &
    \begin{tabular}{l}
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-22-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-77-trim.png}\end{minipage}\\\\
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-64-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-97-trim.png}\end{minipage}
    \end{tabular}\\
    \multicolumn{1}{c|}{}&&\\\cline{2-3}
    \multicolumn{1}{c|}{}&&\\
    \multicolumn{1}{c|}{    \begin{tabular}{c}\textbf{Far apart in }\\ \textbf{image space} \end{tabular}}    &        \begin{tabular}{l}
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-52-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-79-trim.png}\end{minipage}\\\\
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-16-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-17-trim.png}\end{minipage}
    \end{tabular}&
    \begin{tabular}{l}
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-75-extra-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-89-trim.png}\end{minipage}\\\\
      \begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-3-trim.png}\end{minipage}\begin{minipage}{\similarityArrowSize}$\leftrightarrow$\end{minipage}\begin{minipage}{\similaritySize}\includegraphics[width = \similaritySize]{figures/expert-39-trim.png}\end{minipage}
    \end{tabular}
\\    \multicolumn{1}{c|}{}&&\\\cline{2-3}\cline{2-3}    
  \end{tabular}

  \centering
\caption{Pairs of images either close together or far apart in different features spaces.  The symbol $\leftrightarrow$ points to the compared images. Features of the program capture abstract notions like symmetry and repetition. Distance metric over images is $L_{\text{learned}}(\cdot|\cdot)$ (see Section~\ref{generalizingTheHandDrawings}). The off-diagonal entries highlight the difference between these metrics: similarity of programs captures high-level features like repetition and symmetry, whereas similarity of images correspondends to similar drawing commands being in similar places. See Section 2 of the supplement for more examples.}\label{similarityMatrix}  \end{figure}

@@

\subsection{Extrapolating figures}
Having access to the source code of a graphics program facilitates coherent, high-level edits to the figure generated by that program. 
For example,
we could change all of the circles to squares or make all of the lines be dashed.
We extrapolate figures
by increasing the number of times that loops are executed.
Extrapolating repetitive visuals patterns comes naturally to humans,
and building this ability into an application is practical:
imagine hand drawing a repetitive graphical model structure
and having our system automatically induce and extend the pattern.
Fig.~\ref{extrapolationFigure} shows extrapolations produced by our system.
\begin{figure}[H]
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix1.png}
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix2.png}
  \includegraphics[width = \textwidth]{figures/extrapolationMatrix3.png}  
  \caption{Top, white: hand drawings. Bottom, black: extrapolations produced by
running loops for extra iterations.}\label{extrapolationFigure}
  \end{figure}
%\input{extrapolations.tex}
